Lossy compression of DNA sequencing quality data

### 2015-05-12 manuscript/q-compression.tex
a few more small changes

As the cost of sequencing continues to decrease, the rate of sequence
data production is increasing, placing greater demands on storing and
transferring these vast amount of data. Most methods of sequencing
data compression focus on compressing nucleotide information without
any loss of information.  Quality data, however, have different
properties than nucleotide data, and methods compressing nucleotide
sequences efficiently do not perform as well on quality
sequences. Furthermore, while lossless representation is necessary for
nucleotide sequences, it is not an essential requirement for quality
values.

@@

While we focus only on two bins in this work, a greater number of bins
can be used in practice. For example, an additional third bin (good,
bad, and intermediate) can be used for base qualities near the border
between bins in the 2-bin encoding. Examining the trade-off of
additional bins and their resulting compressibility and effect on
downstream analyses are not considered here.

@@

It is important to highlight that using uncompressed reads does not
produce the best assembly in terms of any of the metrics. The
uncompressed reads score worse than the top overall assembly
(256-profile encoding) in terms of assembled bases, missing reference
bases, N50, \textsc{snp}s, indels $>$5bp, and relocations. The
assembly using the original uncompressed reads has an error rate of
$\sim$8.75 errors/100 kb of assembled sequence, while the 256-profile
encoding has an error rate of $\sim$8.02 errors/100 kbp (Supplemental
Table 3).

>>>>

As the cost of sequencing continues to decrease, the rate of sequence
data production is increasing, placing greater demands on storing and
transferring these vast amount of data. Most methods of sequencing
data compression focus on compressing nucleotide information without
any loss of information.  Quality data, however, have different
properties than nucleotide data, and methods compressing nucleotide
sequences efficiently do not perform as well on quality
sequences. Furthermore, although lossless representation is necessary
for nucleotide sequences, it is not an essential requirement for
quality values.

@@

Although we focus only on two bins in this work, a greater number of
bins can be used in practice, albeit with some higher cost in memory.
For example, an additional third bin, thus giving bins of good, bad,
and intermediate, could be used for base qualities near the border
between bins in the 2-bin encoding. Examining the trade-off of
additional bins and their resulting compressibility and effect on
downstream analyses are not considered here.

@@

It is important to highlight that using uncompressed reads does not
produce the best assembly in terms of any of the metrics. The
uncompressed reads score worse than the top overall assembly
(256-profile encoding) in terms of assembled bases, missing reference
bases, N50, \textsc{snp}s, indels $>$5bp, and relocations. The
assembly using the original uncompressed reads has an error rate of
$\sim$8.75 errors/100 kb of assembled sequence, and the 256-profile
encoding has an error rate of $\sim$8.02 errors/100 kbp (Supplemental
Table 3).
